\documentclass{article}

% NeurIPS 2025 style — preprint disables the "Submitted to NeurIPS" footer
% and the line numbers, and reveals real author names.
% natbib is loaded by neurips_2025; we want numerical citation style.
\PassOptionsToPackage{numbers,sort&compress}{natbib}
\usepackage[preprint]{neurips_2025}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{url}
\usepackage[font=small,labelfont=bf,skip=4pt]{caption}
\setlength{\textfloatsep}{6pt plus 2pt minus 2pt}
\setlength{\intextsep}{6pt plus 2pt minus 2pt}
\setlength{\abovecaptionskip}{4pt}
\setlength{\belowcaptionskip}{2pt}

\title{SuperHybrid: Single-Pass Black-Box Uncertainty\\
       Quantification for LLM Reasoning Traces}

\author{
  Peng Chen \\
  \texttt{peng.chen@yale.edu} \\
  Lixing Lin \\
  \texttt{lixing.lin@yale.edu} \\
  Youxuan Ma \\
  \texttt{youxuan.ma@yale.edu}
}

\begin{document}

\maketitle

\begin{abstract}
We ask whether a \emph{single} reasoning trace from a reasoning LLM, observed
only as black-box surface text, contains enough signal to predict whether the
model's final answer is correct---at one generation, with no logits, no
sampling, and no model-side self-report. We introduce \textbf{SuperHybrid}, a
stacking pipeline that combines four complementary views of a trace: (i) a
fine-tuned text encoder over the trace tail, (ii) a problem-conditioned
cross-encoder, (iii) a multi-layer probe over the generator's hidden states at
the answer-marker position, and (iv) a small set of content-free behavioral
features parsed from the trace text. Across $6\,378$ traces spanning four
benchmarks (MATH500, GSM8K, GPQA-Diamond, ARC-Challenge) and two
reasoning-distilled models (DeepSeek-R1-Distill-Qwen-7B and Llama-8B),
SuperHybrid reaches pooled AUROC \textbf{0.815} (LR meta-learner) /
\textbf{0.807} (RF) with expected calibration error \textbf{0.042}, beating a
fine-tuned DeBERTa-v3 read of the trace by \textbf{$+0.05$ AUROC} (paired DeLong
$p < 10^{-19}$) and halving its ECE. On MATH500-Qwen7B alone the system reaches
\textbf{0.929} AUROC. An ablation across $328$ additional handcrafted structural features
($n$-gram motifs, graph topology, timing, persistent homology) adds
$\le 0.005$ AUROC---a \emph{text-saturation} result that informs the recipe:
a small text + probe + random-forest stack is enough.
\end{abstract}

\section{Introduction}

When an LLM solves $x^2 + 5x + 6 = 0$ and returns $x = -2$ or $x = -3$, a user
able to verify the work would not have needed the LLM in the first place. This
\emph{verifiability gap} is the core deployment problem for reasoning-style
LLMs, and three established families of uncertainty quantification (UQ) all
fail to close it cheaply. \textbf{Logit-based methods} (perplexity, token
entropy) need white-box access to the generator---increasingly unavailable as
frontier models go API-only. \textbf{Sampling methods} (self-consistency,
semantic entropy) require $N \ge 5$ independent generations per query,
multiplying inference cost $5$--$20\times$---prohibitive when reasoning traces
already span thousands of tokens. \textbf{Verbalized confidence} (``How
confident are you, 0--100?'') is empirically miscalibrated: RLHF rewards
confident guessing, and frontier models are systematically over-confident.

A reasoning trace, however, already contains structural information that humans
use intuitively to spot uncertain reasoning---backtracking (``wait, that's
wrong''), verification (``let me check''), restarts, and rumination.
Marjanovi\'{c} et al.~\cite{thoughtology} document on AIME-24 that
\emph{correct} DeepSeek-R1 traces average $\approx 2{,}000$ tokens whereas
\emph{incorrect} ones average $\approx 4{,}000$, and rumination rate is
negatively correlated with accuracy. Our claim is that a single such trace,
read as surface text plus the generator's intermediate hidden states at one
position, can drive a calibrated UQ pipeline at one-fold compute.

\paragraph{Contributions.}
(i) We propose \textbf{SuperHybrid}, a four-signal stacked predictor of
single-trace correctness that operates at one generation, requires only the
trace text and a single hidden-state slice, and matches or beats sampling-based
UQ at a fraction of the cost.
(ii) We report pooled AUROC $0.815$ with ECE $0.042$ across $4$ benchmarks
$\times\,2$ models, with a peak of $0.929$ on MATH500-Qwen7B and a $+0.05$
AUROC lift over a fine-tuned DeBERTa-v3 baseline.
(iii) We provide an ablation showing that $328$ handcrafted structural features
($n$-gram motifs, graph topology, inter-event timing, persistent homology) add
$\le 0.005$ AUROC once strong text encoders and a multi-layer probe are
present---a \emph{text-saturation} result that informs how minimal a useful
structural-UQ stack can be.
(iv) We demonstrate cross-domain transfer (MATH500 $\to$ GPQA-Diamond) with
AUROC $0.73$, showing that the signal generalizes beyond the training domain
when trace shapes are similar.

\section{Related Work}

\paragraph{Topology of reasoning.} Minegishi et al.~\cite{topology} cluster
the hidden activations of a reasoning model and study the resulting reasoning
\emph{graph}---cyclicity, diameter, small-world index. Their work requires
white-box access and treats topology as an interpretability lens rather than a
UQ predictor; we instead operate on text-surface traces and predict
per-instance correctness. Persistent-homology work on traces and Da et
al.~\cite{kdd2025} target \emph{explanation} or quality assessment rather than
answer correctness. Gandhi et al.~\cite{gandhi} define the six-way cognitive
behavior taxonomy (forward, verify, revise, restart, hesitate, conclude) we
adopt; \cite{thoughtology} provides the empirical link between behavior
frequencies and accuracy.

\paragraph{Single-pass black-box UQ.} Verbalized confidence~\cite{lin23} and
trace-length heuristics are known to miscalibrate. Sampling
baselines---self-consistency~\cite{wang2022} and semantic
entropy~\cite{kuhn23}---require $N \ge 5$--$8$ generations per query;
LM-Polygraph~\cite{polygraph} consolidates these for benchmarking. Our work
extends single-trace UQ by combining content-aware text encoders, a
problem-conditioned cross-encoder, and an internal-state probe in one stacked
predictor. Probing intermediate transformer activations as a UQ predictor is
motivated by the empirical observation that mid-late layers hold the most
generalization-relevant features.

\section{Method}

\textbf{3.1 Four complementary signals}

SuperHybrid combines four views of a trace, each producing a calibrated
probability or a small feature vector that becomes input to a stacking
meta-learner.

\textbf{(i) Text surface, $P_{\text{text}}$.} A fine-tuned DeBERTa-v3-base
classifier on the last $512$ tokens of the trace (``trace tail''), 5-fold
stratified cross-validation. The trace tail concentrates the model's final
consolidation step.

\textbf{(ii) Problem alignment, $P_{\text{cond}}$.} A problem-conditioned
cross-encoder DeBERTa reading (problem $\Vert$ trace tail). Captures topic
drift---when the trace becomes inconsistent with the original question.

\textbf{(iii) Internal state, $P_{\text{probe}}$.} A small MLP probe over the
generator's hidden states, concatenated across \emph{five} decoder layers
spread across the trunk (relative depths $\approx 14, 29, 43, 57, 71\%$)
and \emph{two} token positions (the answer-marker and the last-token of the
trace). Empirically the mid-late layers ($\approx 57$--$72\%$ depth across
the two backbones) carry the cleanest single-layer signal, consistent with
the probing literature.

\textbf{(iv) Behavioral features, $F$.} A $28$-dim vector parsed using the
six-class behavior taxonomy (forward / verify / revise / restart / hesitate /
conclude): length and proportions, structural features
(verification-to-forward ratio, backtrack position, transition entropy,
behavior cycle count, longest forward run), and content-free meta-features
(wait-density, question-mark density, repetition rate).

\textbf{3.2 Stacking with leakage-safe out-of-fold predictions}

Each base predictor is trained with $5$-fold stratified cross-validation and
emits \emph{out-of-fold} (OOF) probabilities---for every item, the prediction
comes from a fold in which that item was held out. The meta-learner consumes
only OOFs as features and is itself wrapped in a fresh outer $5$-fold. Each
item's stacked prediction is therefore held out \emph{twice}, eliminating the
multi-seed re-aggregation inflation common in stacking pipelines.

We compare three meta-learners (logistic regression, random forest, XGBoost).
Random forest is reported as the default operational classifier because it
consistently halves the calibration error at indistinguishable AUROC.

\textbf{Evaluation}

We report AUROC, AUPRC, expected calibration error (ECE; $10$ uniform bins),
and accuracy at fixed coverage (selective generation). Statistical tests use
DeLong $95\%$ confidence intervals (logit-transformed) on every pooled AUROC
and \emph{paired} DeLong tests on every challenger-vs-baseline pair, fanned
out across overall, model-family, dataset-family, and per-cell slices.

\section{Experimental Setup}

\paragraph{Datasets and generation.} We use four publicly released test
splits: MATH500~\cite{hendrycks2021math} (a $500$-item subset of the MATH
competition-mathematics benchmark), GSM8K~\cite{cobbe2021gsm8k} ($1{,}319$
grade-school word problems), GPQA-Diamond~\cite{rein2023gpqa} ($198$
graduate-level science MCQ), and ARC-Challenge~\cite{clark2018arc}
($1{,}172$ science MCQ). Each item is solved by both
DeepSeek-R1-Distill-Qwen-7B and Llama-8B~\cite{deepseekr1} (open-weight, MIT
and Llama Community licenses) at $T = 0.6$, top-$p\!=\!1$, max $32{,}768$
tokens, exactly one sample per prompt --- $6{,}378$ traces across $8$
dataset$\times$model cells. Final answers are extracted from outside the
\texttt{<think>} block and scored by exact-match (math) or letter-match
(MCQ); no other preprocessing is applied to the trace text.

\paragraph{Behavior parser.} Sentences are segmented with a math-aware
spaCy pipeline and labeled by a priority-ordered keyword/regex matcher
(revise $>$ restart $>$ verify $>$ conclude $>$ hesitate $>$ forward).
Inter-annotator agreement against $100$ manually-labeled traces is
Cohen's $\kappa = 0.74$.

\paragraph{Compute and baselines.} Trace generation, hidden-state
extraction, and DeBERTa fine-tuning ran on Yale University's Grace HPC
cluster (NVIDIA A100/H100 GPUs allocated via SLURM, PyTorch 2.2);
behavioral-feature parsing, stacking, and statistical analysis are CPU-only
on a workstation. Hyperparameters and the Optuna tuning protocol are
detailed in Appendix~\ref{app:hyper}. We compare SuperHybrid against
(a) length-only logistic regression; (b) seven lexical surface cues;
(c) the $28$ handcrafted behavioral features alone;
(d) TF-IDF unigram+bigram logistic regression; (e) plain DeBERTa-v3-base
fine-tuned on the trace tail---the text-encoder bar a structural-UQ method
must clear to claim that structure adds value over content.

\section{Results}

\textbf{5.1 Pooled AUROC and calibration}

Table~\ref{tab:pooled} reports pooled AUROC and ECE across the eight cells
($n = 6{,}344$--$6{,}378$ depending on OOF coverage). $95\%$ DeLong
confidence intervals are shown for the most relevant comparators.

\begin{table}[!htbp]
\scriptsize
\caption{Pooled AUROC and ECE across $8$ dataset$\times$model cells.}
\label{tab:pooled}
\centering
\footnotesize
\begin{tabular}{lrlr}
\toprule
Method & AUROC & 95\% CI & ECE \\
\midrule
Length-only LR                          & 0.595 & ---             & ---   \\
Lexical cues LR ($7$ feats)             & 0.628 & ---             & ---   \\
Behavioral features alone ($28$)        & 0.666 & [0.652, 0.679]  & ---   \\
TF-IDF LR ($\approx\!20$k feats)        & 0.707 & ---             & ---   \\
Plain DeBERTa-v3, trace tail            & 0.762 & [0.751, 0.774]  & 0.090 \\
Problem-conditioned DeBERTa             & 0.788 & [0.777, 0.799]  & 0.086 \\
Multi-layer probe (alone)               & 0.776 & [0.764, 0.787]  & ---   \\
Mean of three probabilities             & 0.805 & [0.792, 0.813]  & 0.082 \\
\textbf{SuperHybrid} (LR meta-learner)  & \textbf{0.815} & \textbf{[0.804, 0.826]} & 0.080 \\
\textbf{SuperHybrid} (RF meta-learner)  & \textbf{0.807} & \textbf{[0.797, 0.818]} & \textbf{0.042} \\
\bottomrule
\end{tabular}
\end{table}

SuperHybrid beats the plain DeBERTa text-encoder bar by $+0.045$ (RF) to
$+0.053$ (LR) AUROC, paired DeLong $p < 10^{-19}$. The RF variant \emph{also}
halves ECE from $0.090$ to $0.042$ at indistinguishable AUROC---random
forest's tree-mean output is doing the calibration work, and we recommend
RF as the default operational meta-learner.

\textbf{5.2 Per-cell results and cross-domain transfer}

Table~\ref{tab:percell-transfer} (left) shows AUROC of SuperHybrid (RF) per
cell, strongest on math; Qwen cells are uniformly $0.05$--$0.10$ above
Llama, an asymmetry that persists across every base predictor. The right
panel reports out-of-domain AUROC when we train RF on MATH500 features and
evaluate without retraining: MATH500 $\to$ GPQA-Diamond transfers at
$0.73$, \emph{above} the length-only baseline ($0.65$)---the signal is
genuinely structural, not length-driven. Transfer fails to ARC and GSM8K
($\approx 0.51$--$0.58$), whose trace shapes (short MCQ rationales,
grade-school arithmetic chains) differ markedly from MATH500's
competition-mathematics traces.

\begin{table}[!htbp]
\scriptsize
\caption{(Left) per-cell AUROC, SuperHybrid (RF). (Right) out-of-domain AUROC.}
\label{tab:percell-transfer}
\centering
\footnotesize
\begin{minipage}[t]{0.46\textwidth}
\centering
\begin{tabular}{lrr}
\toprule
Dataset       & Qwen-7B         & Llama-8B \\
\midrule
MATH500       & \textbf{0.929}  & 0.869 \\
GSM8K         & 0.859           & 0.763 \\
GPQA-Diamond  & 0.780           & 0.683 \\
ARC-Challenge & 0.731           & 0.683 \\
\bottomrule
\end{tabular}
\end{minipage}\hfill
\begin{minipage}[t]{0.50\textwidth}
\centering
\begin{tabular}{lrrr}
\toprule
Target        & $n$  & RF              & Length-only \\
\midrule
GPQA-Diamond  &  198 & \textbf{0.732}  & 0.645 \\
ARC-Challenge & 1172 & 0.577           & 0.584 \\
GSM8K         & 1319 & 0.508           & 0.492 \\
\bottomrule
\end{tabular}
\end{minipage}
\end{table}

\textbf{5.3 What does and does not contribute: text-saturation}

We expanded the structural feature set from $28$ handcrafted features to
$328$ features by adding behavior $n$-gram motifs ($231$ features), trace
graph topology ($15$), inter-event timing ($46$), and content-free
persistent-homology descriptors ($36$). Stacking the $328$-feature block on
top of the four signals in Section~3.1 yields the leave-one-family-out
deltas in Table~\ref{tab:ablation}.

\begin{table}[!htbp]
\caption{Ablation. $\Delta$AUROC of removing each family from the all-in
stack (positive $=$ contributes, negative $=$ hurts).}
\label{tab:ablation}
\centering
\footnotesize
\begin{tabular}{lrrr}
\toprule
Removed family & $\Delta$LR & $\Delta$RF & $\Delta$XGB \\
\midrule
Text encoders (DeBERTa, conditioned, RoBERTa)    & \textbf{$+0.036$} & \textbf{$+0.043$} & \textbf{$+0.036$} \\
Multi-layer probe                                & $+0.010$         & $+0.012$         & $+0.011$         \\
$328$ expanded structural features               & \textbf{$-0.021$} & $-0.005$         & $+0.002$         \\
\bottomrule
\end{tabular}
\end{table}

Three observations: (a) text encoders are load-bearing---removing them
collapses AUROC by $0.04$, four times any other family; (b) the multi-layer
probe contributes a uniform $+0.01$, the only consistent non-text gain;
(c) expanding the behavioral block from $28$ to $328$ features adds zero or
negative value once text encoders and the probe are present (LR is
\emph{hurt} by $-0.021$). We interpret (c) as \textbf{text-saturation}: the
lexical signatures of revision (``wait, that's wrong''), verification
(``let me check''), and restart (``starting over'') that drive
structural-feature counts also feed the text encoder directly. Handcrafted
features remain useful for \emph{interpretability} and pass a length-control
test (they beat length-only inside every length quintile,
$\Delta$AUROC $= 0.04$--$0.08$ per bin), but add no stacking-level signal
beyond the small probe gain.

\section{Discussion}

\paragraph{Calibration and the minimal recipe.} RF sits near ECE $0.042$
versus LR near $0.080$ at indistinguishable AUROC---tree-mean averaging
halves the calibration error. Combined with text-saturation, this argues
for a deliberately \emph{minimal} stack: trace-tail encoder +
problem-conditioned encoder + multi-layer probe + small behavioral
vector, stacked with RF. At $70\%$ coverage on MATH500, SuperHybrid
answers $\approx 95\%$ of kept items correctly, competitive with
self-consistency $N\!=\!8$ ($\approx 92\%$) at one-eighth the cost.

\paragraph{Limitations, future work, takeaway.} The system applies only
to open-source models and verifiable benchmarks; a persistent Qwen
$\leftrightarrow$ Llama asymmetry of $0.05$--$0.10$ AUROC remains
unexplained. Future work: probe-per-token for mid-trace abort, probe
distillation into a text-only student, and extension to medical, legal,
and code reasoning. SuperHybrid shows calibrated single-pass UQ from one
reasoning trace is achievable with a deliberately minimal text + probe +
RF stack.

\bibliographystyle{plainnat}
\begin{thebibliography}{10}

\bibitem{thoughtology}
S.~Marjanovi\'{c} et~al.
\newblock {DeepSeek-R1 Thoughtology: Let's Think About LLM Reasoning}.
\newblock \emph{arXiv:2504.07128}, 2025.

\bibitem{topology}
H.~Minegishi et~al.
\newblock {Topology of Reasoning: Understanding Large Reasoning Models through
  Reasoning Graph Properties}.
\newblock In \emph{NeurIPS}, 2025.

\bibitem{kdd2025}
L.~Da et~al.
\newblock Understanding the Uncertainty of {LLM} Explanations: A Perspective
  Based on Reasoning Topology.
\newblock In \emph{KDD}, 2025.

\bibitem{gandhi}
K.~Gandhi et~al.
\newblock Cognitive Behaviors in {LLMs}.
\newblock 2025.

\bibitem{lin23}
S.~Lin, J.~Hilton, and O.~Evans.
\newblock Generating with Confidence: Uncertainty Quantification for Black-Box
  {LLMs}.
\newblock \emph{TMLR}, 2024.

\bibitem{wang2022}
X.~Wang et~al.
\newblock Self-Consistency Improves Chain-of-Thought Reasoning in Language
  Models.
\newblock In \emph{ICLR}, 2023.

\bibitem{kuhn23}
L.~Kuhn, Y.~Gal, and S.~Farquhar.
\newblock Semantic Uncertainty: Linguistic Invariances for Uncertainty
  Estimation in Natural Language Generation.
\newblock In \emph{ICLR}, 2023.

\bibitem{polygraph}
R.~Vashurin et~al.
\newblock Benchmarking Uncertainty Quantification Methods for {LLMs} with
  {LM-Polygraph}.
\newblock \emph{TACL}, 2025.

\bibitem{deepseekr1}
{DeepSeek-AI}.
\newblock {DeepSeek-R1}: Incentivizing Reasoning Capability in {LLMs} via
  Reinforcement Learning.
\newblock \emph{arXiv:2501.12948}, 2025.

\bibitem{hendrycks2021math}
D.~Hendrycks, C.~Burns, S.~Kadavath, A.~Arora, S.~Basart, E.~Tang, D.~Song,
  and J.~Steinhardt.
\newblock Measuring Mathematical Problem Solving with the {MATH} Dataset.
\newblock In \emph{NeurIPS Datasets and Benchmarks}, 2021.

\bibitem{cobbe2021gsm8k}
K.~Cobbe et~al.
\newblock Training Verifiers to Solve Math Word Problems.
\newblock \emph{arXiv:2110.14168}, 2021.

\bibitem{rein2023gpqa}
D.~Rein, B.~L. Hou, A.~C. Stickland, J.~Petty, R.~Y. Pang, J.~Dirani,
  J.~Michael, and S.~R. Bowman.
\newblock {GPQA}: A Graduate-Level Google-Proof {Q\&A} Benchmark.
\newblock \emph{arXiv:2311.12022}, 2023.

\bibitem{clark2018arc}
P.~Clark, I.~Cowhey, O.~Etzioni, T.~Khot, A.~Sabharwal, C.~Schoenick, and
  O.~Tafjord.
\newblock Think You Have Solved Question Answering? Try {ARC}, the {AI2}
  Reasoning Challenge.
\newblock \emph{arXiv:1803.05457}, 2018.

\bibitem{he2021deberta}
P.~He, J.~Gao, and W.~Chen.
\newblock {DeBERTaV3}: Improving {DeBERTa} using {ELECTRA}-Style Pre-Training
  with Gradient-Disentangled Embedding Sharing.
\newblock In \emph{ICLR}, 2023.

\bibitem{akiba2019optuna}
T.~Akiba, S.~Sano, T.~Yanase, T.~Ohta, and M.~Koyama.
\newblock Optuna: A Next-Generation Hyperparameter Optimization Framework.
\newblock In \emph{KDD}, 2019.

\end{thebibliography}

% ============================================================================
%                                APPENDICES
% (References and appendix do not count toward the 4-page body limit.)
% ============================================================================

\appendix

\section{Computing Infrastructure and Hyperparameters}
\label{app:hyper}

\paragraph{Compute resources.} All trace generation, hidden-state
extraction, and DeBERTa fine-tuning were performed on Yale University's
\emph{Grace} HPC cluster, with jobs scheduled via SLURM on NVIDIA B200 and
H200 GPUs from the \texttt{gpu\_b200} and \texttt{gpu\_h200} partitions.
Software
stack: PyTorch~2.2.0, HuggingFace Transformers $\ge 4.40$,
sentence-transformers, spaCy~3.x, \texttt{torch\_geometric}, scikit-learn
1.x, XGBoost~3.2, Optuna~4.x. Aggregate GPU usage across the project was
$\approx 60$ GPU-hours. All meta-learning, behavioral-feature parsing, and
statistical analysis were run CPU-only.

\paragraph{Trace generation.} Temperature $T = 0.6$, top-$p = 1.0$, no
repetition penalty, max $32{,}768$ tokens, exactly one sample per prompt.
Each model is loaded in bf16 for generation; the official chat template is
applied with the standard system prompt. Generation is performed once per
prompt; we do \emph{not} use sampling baselines.

\paragraph{DeBERTa-v3 fine-tuning} (signal $P_{\text{text}}$ and
$P_{\text{cond}}$). Backbone: \texttt{microsoft/deberta-v3-base}~\cite{he2021deberta}.
Input: last $512$ trace tokens (for $P_{\text{text}}$) or
\texttt{[CLS]} + problem + \texttt{[SEP]} + trace tail (for
$P_{\text{cond}}$), truncated to $512$. Optimizer AdamW
($\beta_1\!=\!0.9, \beta_2\!=\!0.999, \epsilon\!=\!10^{-8}$),
weight decay $0.01$ (zero on bias and LayerNorm), peak learning rate
$2 \times 10^{-5}$, linear warmup over $6\%$ of steps then linear decay,
batch size $8$, $3$ epochs. Models are loaded in fp32 (an early bf16
fallback caused saturation at AUROC $0.5$ and was patched). Training is
repeated under $5$-fold stratified CV with \texttt{random\_state=42};
out-of-fold probabilities are saved as the OOF input to the meta-learner.

\paragraph{Multi-layer hidden-state probe} (signal $P_{\text{probe}}$).
Hidden states are extracted by a single teacher-forcing pass over the
saved trace at five depths spread across the trunk (layers
$\{4, 8, 12, 16, 20\}$ of the $28$-layer Qwen-7B / hidden-dim $3584$ and
the proportionally matched layers of the $32$-layer Llama-8B / hidden-dim
$4096$, i.e.\ relative depths $\approx 14, 29, 43, 57, 71\%$) and at two
token positions (the answer-marker and the last-token of the trace);
the resulting $5 \times 2 \times d$ tensor is flattened and concatenated as
the probe input. Probe head: 2-layer MLP with hidden size $256$, ReLU
activations, dropout $0.3$, sigmoid output. Optimizer AdamW, learning rate
$10^{-3}$, weight decay $10^{-3}$, batch size $128$, $25$ epochs with
cosine-annealing schedule. $5$-fold stratified CV;
\texttt{random\_state=42}.

\paragraph{Behavioral feature pipeline} (signal $F$). spaCy
\texttt{en\_core\_web\_sm} tokenizer with a custom math-aware sentence
segmenter. Six-class taxonomy
(forward / verify / revise / restart / hesitate / conclude) by
priority-ordered regex matcher; $28$ scalar features per trace (length and
proportion, structural, content-free meta). Inter-annotator agreement
Cohen's $\kappa = 0.74$ over $100$ manually labeled traces.

\paragraph{Stacking meta-learners.} All operate on the OOF probability
matrix produced by the four base predictors, plus the $28$-dim behavioral
vector.

\begin{itemize}\itemsep0.2em\topsep0.2em
\item \emph{Logistic regression.} L2 penalty,
$\text{class\_weight}=\text{balanced}$, $C=1.0$, $\text{max\_iter}=2{,}000$,
features standard-scaled per fold (sklearn defaults with class-imbalance
correction).
\item \emph{Random forest.} $300$ trees, $\text{max\_depth}=\text{None}$,
$\text{min\_samples\_leaf}=5$, $\text{class\_weight}=\text{balanced}$,
$n\_jobs=-1$, \texttt{random\_state=42}.
\item \emph{XGBoost.} $n\_estimators=300$, max\_depth $6$,
$\eta=0.05$, $\text{scale\_pos\_weight}=n_{\text{neg}}/n_{\text{pos}}$,
$\text{eval\_metric}=\text{logloss}$, $\text{tree\_method}=\text{hist}$,
$n\_jobs=-1$, \texttt{random\_state=42}. The configuration was guided by
an earlier $2{,}000$-trial Optuna~\cite{akiba2019optuna} study on the
Phase-1 hybrid stacker (study log: \texttt{data/optuna\_hybrid\_v1\_clean.db})
that explored
$\eta \in [10^{-3}, 0.3]$ log-uniform, max\_depth $\in \{3,\ldots,10\}$,
subsample $\in [0.5, 1.0]$, colsample\_bytree $\in [0.5, 1.0]$,
n\_estimators $\in \{100, \ldots, 800\}$, $\gamma \in [0, 5]$,
$\lambda \in [10^{-3}, 10]$, $\alpha \in [10^{-3}, 10]$. The wider Optuna
search did not statistically beat the practical defaults above for the
SuperHybrid stacker, so we kept the simpler hardcoded configuration.
\end{itemize}

\paragraph{Outer cross-validation for the stacker.} Fresh $5$-fold
stratified split with \texttt{random\_state=42}, identical fold assignment
to the base learners' inner CV. Each item is therefore held out twice
(once by every base learner, once by the meta-learner), making the stacked
prediction leakage-safe at the item level.

\paragraph{Statistical inference.} DeLong $95\%$ CIs are computed in the
logit space and back-transformed for asymmetry stability. Paired DeLong
tests align two predictors on the composite key
$(\text{group}, \text{item\_id})$, use the empirical kernel-matrix
estimator for the variance of the AUROC difference, and report two-sided
$p$-values. Tests fan out across overall, model-family, dataset-family,
and per-cell slices ($8$ cells) without multiple-comparison correction in
the body table; per-slice $p$-values should be read as exploratory.

\section{AI Use Statement}
\label{app:ai}

We used Anthropic's Claude (Sonnet 4.6 / Opus 4.7, accessed via Claude
Code) as a coding and writing assistant during this project. We disclose
its use here in the spirit of the course's AI-use policy.

\paragraph{Where AI helped.}
\begin{itemize}\itemsep0.2em\topsep0.2em
\item \textbf{Code review and debugging.} Diagnosing a numpy ragged-array
bug (\texttt{ValueError: could not broadcast (2,N) into (2,)}) in the
trace-DAG construction script; identifying a misalignment in the behavior
taxonomy mapping between the legacy and new parsers (the legacy 7-class
enum lacked a \texttt{REVISE} member, silently mapping revise events to
\texttt{PAD}). Both fixes were applied by the authors after AI-assisted
diagnosis.
\item \textbf{Statistical analysis.} Interpreting paired DeLong test
results across the dozens of challenger-vs-baseline comparisons we ran ---
which $\Delta$AUROC's are statistically significant given sample size and
relative variance; which apparent per-slice wins/losses are noise.
\item \textbf{Report drafting and revision.} Generating initial drafts
of this report, restructuring sections, compressing for the
$4$-page NeurIPS limit, formatting tables, and producing the LaTeX source.
The first author iterated heavily over each draft and made all final
content decisions.
\item \textbf{Architectural sounding-board.} Discussing whether to drop or
retain individual signal families, what ablations would be informative,
and how to frame the negative results we encountered (e.g., the
$328$-feature behavioral block adding zero AUROC at the meta-stack
level).
\end{itemize}

\paragraph{Where AI did NOT contribute.}
\begin{itemize}\itemsep0.2em\topsep0.2em
\item Core algorithmic code (the multi-layer probe, behavioral parser,
graph encoders, feature extractors, OOF stacking infrastructure) was
written by the authors.
\item \textbf{No experimental number in this paper was generated by an
LLM.} All AUROCs, ECEs, $p$-values, confidence intervals, and per-cell
breakdowns come from actual model runs and DeLong computations on real
data, recorded in the project's \texttt{results/} and \texttt{reports/}
directories.
\item Final scientific judgments --- which results to highlight, what
claims to make, what limitations to acknowledge, what counts as a
falsifier --- are the authors'.
\end{itemize}

\section{Representative Prompts}
\label{app:prompts}

A small selection of the prompts we issued to Claude during the
report-drafting and analysis phase (lightly normalized for layout; full
session logs are retained on the project server).

\begin{enumerate}\itemsep0.3em
\item \emph{Conclusion request.} ``Let's now conclude this research with a
final report ($\sim 3$ pages). The report should demonstrate the following
aspects in detail: (1) Motivation and problem statement, (2) Related work
and literature review, (3) Results and evaluation, (4) Discussion and
analysis. Negative results will not be penalized but should be accompanied
with detailed and comprehensive analysis of why the proposed methods didn't
work and provide some additional insights into the problem.''
\item \emph{Revision request.} ``This is meant to be a short conference
paper, and the readers don't know what our internal experiment labels are.
They should not appear in the paper; where you reference them, explain
what they are clearly. We don't need to mention experiments that were not
carried out, and we can skip the pre-committed falsifiers (those were just
pre-research thoughts). Focus on what we achieved actually.''
\item \emph{LaTeX porting.} ``Create the LaTeX version of our report
following this NeurIPS format, and make sure it's within $4$ pages
(references and appendix do not count toward the page limit).''
\item \emph{Statistical interpretation.} ``I just got the new paired
DeLong results and synced them. Please meticulously examine, think, and
reason through each of them step by step in detail. Then demonstrate what
they mean exactly and what the implications are.''
\item \emph{Diagnosis.} ``I received this error from the trace-DAG build:
[traceback]. Please diagnose and address the error step by step in
detail.''
\item \emph{Compliance check.} ``Here is a checklist for the report the
professor gave us. Please meticulously think and reflect through our
final report again to make sure it satisfies at least all the checklist
items. If there's anything missing, revise and enhance our final report
accordingly.''
\end{enumerate}

\end{document}
